Portable Multi- and Many-Core Performance for Finite-Difference/Element Codes; Application to the Free-Surface Component of NEMO (NEMOLite2D 1.0).

### 2017-09-18 docs/papers/nemolite2d_performance/referee_response.txt

Figure 5 - it is a reflection of the fact that Gnu is free while the Intel and
Cray compilers have performance of the generated code as a key part of their
respective business models.

Page 18, line 9 - the NEMOLite2D performance for the 32x32 domain is
significantly greater than for the other domains because it largely
fits within L2 cache. We have calculated that the working-set size for
the momentum kernel is 136*n^2. For 32x32 this gives ~136KB while for
64x64 this gives ~544KB. The former will fit within L2 cache while the
latter will spill to L3 cache. In addition, recall that we only parallelise
the outermost loop (Sec. 2.2.1) so, for the 32x32 domain, there are only
32 items to be shared amongst the threads. We have added a note to the
text regarding this.

Figure 11 - the size of the Momentum kernels meant that
their resource use limited the occupancy that could be achieved.
Therefore COLLAPSE might only help the performance of the other kernels and
they only account for some 30% of the run-time. In addition, the kernels
had been parallelised by enclosing them within a !$acc kernels region
which leaves the details to the compiler. It would seem that in this
case the decisions made by the compiler were good and COLLAPSE provided
little or no benefit. The text has been altered to cover both of these
points.

Page 22, line 11 - we agree with the reviewer's point that there was
no proof that the CUDA version was performing well. However, it is
also not a matter of comparing memory bandwidths on the CPU and GPU -
our work on the roofline model has demonstrated that the Momentum
kernel is not memory-bandwidth bound. Instead, we have added a
paragraph that demonstrates we have used profiling to investigate the
performance limitations on the GPU and are confident that we can do no
better without changing kernel code.

Figure 13 - font size and line widths have been increased. Labels have been
re-arranged to improve clarity.

### 2017-09-18 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: add text about GPU performance quantification

Note that all of the previous optimisations are restricted to the PSy
layer and in most cases are simply a case of adding directives or
clauses to directives. However, the question then arises as to the
cost of restricting optimisations to the PSy layer. In particular, how
does the performance of the OpenACC version of NEMOLite2D compare with
a version where those restrictions are lifted?
Figure~\ref{FIG_gpu_opt_stages} also shows the performance of the
version of the code where the (fused) Momentum kernel has been
replaced by a CUDA kernel. For domains up to $256^2$ the difference in
the performance of the two versions is less than 5\% and for the
larger domains it is at most 12\%. This demonstrates that the
performance cost of the {PS}y{KA}l approach in utilising a GPU for
NEMOLite2D is minimal.  The discrepancy in performance between the
CUDA and OpenACC versions has been fed back to the PGI compiler team.

>>>>

Note that all of the previous optimisations are restricted to the PSy
layer and in most cases are simply a case of adding directives or
clauses to directives. However, the question then arises as to the
cost of restricting optimisations to the PSy layer. In particular, how
does the performance of the OpenACC version of NEMOLite2D compare with
a version where those restrictions are lifted?
Figure~\ref{FIG_gpu_opt_stages} also shows the performance of the
version of the code where the (fused) Momentum kernel has been
replaced by a CUDA kernel. For domains up to $256^2$ the difference in
the performance of the two versions is less than 5\% and for the
larger domains it is at most 12\%.

In order to check the efficiency of the CUDA implementation we
profiled the code on the GPU.  For large problem sizes this
showed that the non-Momentum kernels are memory-bandwidth bound and
account for about 40\% of the runtime. However, the Momentum kernels
were latency limited, getting $\sim$103 GB/s of memory bandwidth. Although
fusing these kernels reduced the required memory bandwidth (and
produced a performance improvement), doing so resulted in a fairly
large kernel requiring a large number of registers. This in turn
reduces the occupancy of the device which exposes memory latencies.

All of this demonstrates that the performance cost of the {PS}y{KA}l
approach in utilising a GPU for NEMOLite2D is minimal.  The
discrepancy in performance between the CUDA and OpenACC versions has
been fed back to the PGI compiler team.
